Classical Planning Algorithms ¡ on the Atari Video Games: Preliminary Results

### 2014-10-31 workshop-paper/experimental_results.tex
Second pass on the Experimental Section

We have tested the $\IW(1)$ and \BFS~described in the previous Section
over $9$ games selected amongst those considered in the \emph{training} set
of \cite{bellemare:jair2013}, and those discussed in \cite{deep-mind-atari}. 
In Table~\ref{performance} the performance of both $\IW(1)$ 
and \BFS, characterized by the average score obtained over all episodes, the standard deviation
of the score obtained and time (in seconds) per lookahead, is compared with that of \UCT, 
\BRFS~and the baseline random action selection. Our experimental 
setup  follows that of \cite{bellemare:jair2013}, with two exceptions. First, rather than
considering $10$ episodes of each game, we have considered just $5$
due to time constraints, as each game can take as much as $40$ hours of 
CPU time. Second, for \UCT, rather than limiting the number of
base policy rollouts to $500$, we have rather limited the number of frames
to be simulated during each lookahead to $100,000$ as \cite{bellemare:jair2013} 
does for \BRFS~and we do for both $\IW(1)$ and \BFS. Other \UCT~parameters such as the exploration constant $C$ and
the planning horizon $m$ have not been modified ($C=0.1$, $m=1500$). We include
the average score reported for \UCT~by \cite{bellemare:jair2013} in Table~\ref{performance}
bracketed with parenthesis for reference. Also unchanged is the discount factor
used by all the algorithms ($\gamma=0.995$) and the rate at which an algorithm is
invoked to select an action (every $5$ frames~\footnote{We note that with this setup, the maximum depth $d$
of the lookahead will be $m/5$, or $300$ nodes.}). All algorithms reuse the tree built between each 
lookahead by setting the root of the tree to the best child of the root in the previous one,
and deleting its siblings and their descendants, as \cite{bellemare:jair2013} does for \UCT. We note that 
this may have a different effect on algorithms where all branches have the same number of
generated nodes, as is the case of \BRFS, than in $\IW(1)$, \BFS~and \UCT, where the depth of
the branches is uneven.

@@

We observe that in $4$ out of $9$ games, $\IW(1)$ outperforms \UCT~limited to the same number
of simulated frames, both in terms of average score attained and average time per call. \BFS~is
generally faster than $\IW(1)$, but only outperforms \UCT~in $2$ out of $9$ games (\Freeway and
\Seaquest) and $\IW(1)$ in $4$ out of $9$ games. Compared with the \UCT~results reported
in \cite{bellemare:jair2013}, that by limiting the number of rollouts allow a $50$\% more frames
to be simulated in each lookahead, $\IW(1)$ still outperforms \UCT~in $3$ out of $9$ games
(\Breakout, \Enduro and \Freeway). As introduced in the previous Section, the order in which
$\IW(1)$ and \BFS~consider actions is randomized, since the novelty of a given state may
change depending on the order in which states are considered. We found this randomization to
have a significant effect in (\textbf{TODO}: What game? \Qbert?) where (\textbf{TODO}: What was
going on?). Last, we note that $\IW(1)$ and \BFS~attain similar performance as the baseline
semi-random algorithm \emph{Perturb} from \cite{bellemare:jair2013}, which beated both \UCT~
and the reinforcement learning algorithms reported.

>>>>

We have tested the $\IW(1)$ and \BFS~described in the previous Section
over $9$ games selected amongst those considered in the \emph{training} set
of \cite{bellemare:jair2013}, and those discussed in \cite{deep-mind-atari}. 
In Table~\ref{performance} the performance of both $\IW(1)$ 
and \BFS, characterized by the average score obtained over all episodes, the standard deviation
of the score obtained and time (in seconds) per lookahead, is compared with that of \UCT, 
\BRFS~and the baseline random action selection. Our experimental 
setup  follows that of \cite{bellemare:jair2013}, with two exceptions. First, rather than
considering $10$ episodes of each game, we have considered just $5$
due to time constraints, as each game can take as much as $40$ hours of 
CPU time. Second, while \UCT~is limited to perform up to $500$ base policy rollouts ($150,000$ frames
are to be simulated while performing the lookahead), both $\IW(1)$ and \BFS are limited to
simulate $100,000$ frames. \UCT~parameters such as the exploration constant $C$ and
the planning horizon $m$ have not been modified ($C=0.1$, $m=300$), the latter being also
used by both $\IW(1)$ and \BFS~to limit the depth of the lookahead. We include
the average score reported for \UCT~by \cite{bellemare:jair2013} in Table~\ref{performance}
bracketed with parenthesis for reference. Also unchanged is the discount factor
used by all the algorithms ($\gamma=0.995$) and the rate at which an algorithm is
invoked to select an action (every $5$ frames). All algorithms reuse the tree built between each 
lookahead by setting the root of the tree to the best child of the root in the previous one,
and deleting its siblings and their descendants, as \cite{bellemare:jair2013} does for \UCT. We note that 
this may have a different effect on algorithms where all branches have the same number of
generated nodes, as is the case of \BRFS, than in $\IW(1)$, \BFS~and \UCT, where the depth of
the branches is uneven.

@@

We observe that in $4$ out of $9$ games, $\IW(1)$ outperforms \UCT~limited to the same number
of simulated frames, both in terms of average score attained and average time per call. \BFS~is
generally faster than $\IW(1)$, but only outperforms \UCT~in $2$ out of $9$ games (\Freeway and
\Seaquest) and $\IW(1)$ in $4$ out of $9$ games. Compared with the \UCT~results reported
in \cite{bellemare:jair2013}, that by limiting the number of rollouts allow a $50$\% more frames
to be simulated in each lookahead, $\IW(1)$ still outperforms \UCT~in $3$ out of $9$ games
(\Breakout, \Enduro and \Freeway). As introduced in the previous Section, the order in which
$\IW(1)$ and \BFS~consider actions is randomized, since the novelty of a given state may
change depending on the order in which states are considered. We found this randomization to
have a significant effect in \Qbert where not randomizing the order in which actions are considered
resulted in $\IW(1)$ being trapped in a loop. Last, we note that both $\IW(1)$ and \BFS~attain similar 
performance on \Freeway as the baseline semi-random algorithm \emph{Perturb} 
from \cite{bellemare:jair2013}, which beated both \UCT~and the reinforcement learning algorithms reported.

### 2014-10-31 workshop-paper/performance_table.tex
Second pass on the Experimental Section

 \caption{\small Random baseline, \BRFS, $\IW(1)$, \BFS, \UCT~performance over 5 runs,
   setting the Atari 2600 random seeds to $0, \ldots, 4$, and maximum episode duration of $18,000$
   frames. \textit{T} is the average CPU T per call to each action selection, \textit{S} is the average score achieved,
   and \textit{SD} is the standard deviation. \UCT~performance
   reported in \cite{bellemare:jair2013} is shown in parenthesis. Numbers
   in bold show best performer in terms of average score. Average score reported in 
   \cite{deep-mind-atari} is reported for reference under the column \Deepmind.
   }
\label{performance}

>>>>

 \caption{\small Random baseline, \BRFS, $\IW(1)$, \BFS, \UCT~performance over 5 runs,
   setting the Atari 2600 random seeds to $0, \ldots, 4$, and maximum episode duration of $18,000$
   frames. \textit{T} is the average CPU time per call to each action selection algorithm, \textit{S} is the average score achieved,
   and \textit{SD} is the standard deviation. \UCT~performance
   reported in \cite{bellemare:jair2013} is shown in parenthesis. Numbers
   in bold show best performer in terms of average score. Average score reported in 
   \cite{deep-mind-atari} is reported for reference under the column \Deepmind.
   }
\label{performance}
